---
title: "Group assessment draft"
author: "TEAM MORSE"
date: "19/08/2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Hi Loc, Darren, and Chloe:
I've looked into the variables closer and tried a few combinations based on what we'd usually look at clinically. I'm happy to answer any clinical concerns (as I'm an icu clinician...), but definitely need everybody's help with the stats/analysis part (...with no maths background)! You may have other ways of looking at the data, so please share. We don't want our patients dying so let's see how we can predict that! :D This is but a start--please feel free to give feedback and continue on!

Cheers,
Dane

```{r}

#load libraries
library(dplyr)
library(ggplot2)

#load ICU data frame
icu_patients_df1 <- readRDS("icu_patients_df1.rds")

#check structure of data frame
str(icu_patients_df1)
head(icu_patients_df1)
summary(icu_patients_df1)



```

total of 2061 observations, 120 variables
mostly numeric, some factors (ICUType, Gender), one logical (status)
lots of variables with NAs
  diasABP, height, weight, 
  survival (since other patients did not die; confirmed by "0" in_hospital_death )

```{r}


# variable names/headings for easy reference
df_variables <- colnames(icu_patients_df1)
df_variables

# univariate exploration for specific variables that may be of importance

glm_ICUtype <- glm(in_hospital_death ~ ICUType, data = icu_patients_df1)
summary(glm_ICUtype)
# ICUtype is a significant variable

glm_Age <- glm(in_hospital_death ~ Age, data = icu_patients_df1)
summary(glm_Age)
# Age is a significant variable

glm_Gender <- glm(in_hospital_death ~ Gender, data = icu_patients_df1)
summary(glm_Gender)
# Gender is not a significant variable

glm_Height <- glm(in_hospital_death ~ Height, data = icu_patients_df1)
summary(glm_Height)
# Height is not a significant variable

glm_BUN_max <- glm(in_hospital_death ~ BUN_max, data = icu_patients_df1)
summary(glm_BUN_max)
# BUN_max is a significant variable

glm_SAPS1 <- glm(in_hospital_death ~ SAPS1, data = icu_patients_df1)
summary(glm_SAPS1)
# SAPS1 is a significant variable

glm_Creatinine_max <- glm(in_hospital_death ~ Creatinine_max, data = icu_patients_df1)
summary(glm_Creatinine_max)
# Creatinine_max is a significant variable

glm_DiasABP_diff <- glm(in_hospital_death ~ DiasABP_diff, data = icu_patients_df1)
summary(glm_DiasABP_diff)
# DiasABP_diff is not a significant variable

glm_FiO2_max <- glm(in_hospital_death ~ FiO2_max, data = icu_patients_df1)
summary(glm_FiO2_max)
# FiO2_max is not a significant variable

glm_GCS_min <- glm(in_hospital_death ~ GCS_min, data = icu_patients_df1)
summary(glm_GCS_min)
# GCS_min is a significant variable

glm_HCO3_diff <- glm(in_hospital_death ~ HCO3_diff, data = icu_patients_df1)
summary(glm_HCO3_diff)
# HCO3_diff is a significant variable

glm_HCT_min <- glm(in_hospital_death ~ HCT_min, data = icu_patients_df1)
summary(glm_HCT_min)
# HCT_min is not a significant variable

glm_HR_diff <- glm(in_hospital_death ~ HR_diff, data = icu_patients_df1)
summary(glm_HR_diff)
# HR_diff is a significant variable

glm_K_diff <- glm(in_hospital_death ~ K_diff, data = icu_patients_df1)
summary(glm_K_diff)
# K_diff is a significant variable

glm_Lactate_max <- glm(in_hospital_death ~ Lactate_max, data = icu_patients_df1)
summary(glm_Lactate_max)
# Lactate_max is a significant variable

glm_MAP_diff <- glm(in_hospital_death ~ MAP_diff, data = icu_patients_df1)
summary(glm_MAP_diff)
# MAP_diff is not a significant variable

glm_Mg_diff <- glm(in_hospital_death ~ Mg_diff, data = icu_patients_df1)
summary(glm_Mg_diff)
# Mg_diff is not a significant variable

glm_Na_diff <- glm(in_hospital_death ~ Na_diff, data = icu_patients_df1)
summary(glm_Na_diff)
# Na_diff is a significant variable

glm_NIDiasABP_diff <- glm(in_hospital_death ~ NIDiasABP_diff, data = icu_patients_df1)
summary(glm_NIDiasABP_diff)
# NIDiasABP_diff is not a significant variable

glm_NIMAP_diff <- glm(in_hospital_death ~ NIMAP_diff, data = icu_patients_df1)
summary(glm_NIMAP_diff)
# NIMAP_diff is not a significant variable

glm_NISysABP_diff <- glm(in_hospital_death ~ NISysABP_diff, data = icu_patients_df1)
summary(glm_NISysABP_diff)
# NISysABP_diff is a significant variable

glm_PaCO2_diff <- glm(in_hospital_death ~ PaCO2_diff, data = icu_patients_df1)
summary(glm_PaCO2_diff)
# PaCO2_diff is a significant variable

glm_PaCO2_max <- glm(in_hospital_death ~ PaCO2_max, data = icu_patients_df1)
summary(glm_PaCO2_max)
# PaCO2_max is not a significant variable

glm_PaO2_max <- glm(in_hospital_death ~ PaO2_max, data = icu_patients_df1)
summary(glm_PaO2_max)
# PaO2_max is a significant variable

glm_Platelets_min <- glm(in_hospital_death ~ Platelets_min, data = icu_patients_df1)
summary(glm_Platelets_min)
# Platelets_min is not a significant variable

glm_RespRate_diff <- glm(in_hospital_death ~ RespRate_diff, data = icu_patients_df1)
summary(glm_RespRate_diff)
# RespRate_diff is a significant variable

glm_SaO2_diff <- glm(in_hospital_death ~ SaO2_diff, data = icu_patients_df1)
summary(glm_SaO2_diff)
# SaO2_diff is not a significant variable

glm_SysABP_diff <- glm(in_hospital_death ~ SysABP_diff, data = icu_patients_df1)
summary(glm_SysABP_diff)
# SysABP_diff is a significant variable

glm_Temp_diff <- glm(in_hospital_death ~ Temp_diff, data = icu_patients_df1)
summary(glm_Temp_diff)
# Temp_diff is a significant variable

glm_TroponinI_max <- glm(in_hospital_death ~ TroponinI_max, data = icu_patients_df1)
summary(glm_TroponinI_max)
# TroponinI_max is not a significant variable

glm_TroponinT_max <- glm(in_hospital_death ~ TroponinT_max, data = icu_patients_df1)
summary(glm_TroponinT_max)
# TroponinT_max is a significant variable

glm_Urine_diff <- glm(in_hospital_death ~ Urine_diff, data = icu_patients_df1)
summary(glm_Urine_diff)
# Urine_diff is a significant variable

glm_WBC_diff <- glm(in_hospital_death ~ WBC_diff, data = icu_patients_df1)
summary(glm_WBC_diff)
# WBC_diff is a significant variable

glm_Weight_max <- glm(in_hospital_death ~ Weight_max, data = icu_patients_df1)
summary(glm_Weight_max)
# Weight_max is a significant variable


# EXPLORING VARIABLES FOR GLM MODEL

# trial 1: after checking the data, variables of interest are chosen for initial analysis
icu_glm_initial_1 <- glm(in_hospital_death ~ 
                     # patient descriptor variables
                     ICUType +
                     Age +
                     Gender +
                     Height +
                     BUN_max +
                     # clinical variables
                     SAPS1 + 
                     Creatinine_max +
                     DiasABP_diff +
                     FiO2_max + 
                     GCS_min +
                     HCO3_diff +
                     HCT_min + 
                     HR_diff + 
                     K_diff + 
                     Lactate_max +
                     MAP_diff + 
                     Mg_diff + 
                     Na_diff +
                     NIDiasABP_diff + 
                     NIMAP_diff +
                     NISysABP_diff +
                     PaCO2_diff +
                     PaCO2_max + 
                     PaO2_max +
                     Platelets_min +
                     RespRate_diff + 
                     SaO2_diff + 
                     SysABP_diff +
                     Temp_diff +
                     TroponinI_max +
                     TroponinT_max +
                     Urine_diff +
                     WBC_diff +
                     Weight_max, data = icu_patients_df1, family = binomial)
summary(icu_glm_initial_1)

'
From this (icu_glm_initial_1), the significant variables were:
BUN_max                              ***
SAPS1                                *  
HCT_min                              ** 
Lactate_max                          .  
PaO2_max                             .  

BUN_max and HCT_min appear come out significant for this initial model. 
SAPS1 is also significant, would like to also compare with SOFA in next model. 
Id explore other variables for the clinical markers lactate, MAP, NISysABP, and PaO2_max
to test for stronger significance (i.e. min, max, or diff)
'

```



```{r}


# trial 2: exploring other variables
# trying other markers for lactate, MAP, NISysABP, and PaO2_max
icu_glm_initial_2 <- glm(in_hospital_death ~ 
                           # patient descriptor variables
                           ICUType +
                           Age +
                           Gender +
                           Height +
                           BUN_max +
                           # clinical variables
                           SAPS1 + 
                           Creatinine_max +
                           DiasABP_diff +
                           FiO2_max + 
                           GCS_min +
                           HCO3_diff +
                           HCT_min + 
                           HR_diff + 
                           K_diff + 
                           Lactate_max +
                           MAP_diff + 
                           Mg_diff + 
                           Na_diff +
                           NIDiasABP_diff + 
                           NIMAP_diff +
                           NISysABP_diff +
                           PaCO2_diff +
                           PaCO2_max + 
                           PaO2_max +
                           Platelets_min +
                           RespRate_diff + 
                           SaO2_diff + 
                           SysABP_diff +
                           Temp_diff +
                           TroponinI_max +
                           TroponinT_max +
                           Urine_diff +
                           WBC_diff +
                           Weight_max +
                           #EXPLORING OTHER VARIABLES
                           SOFA +
                           Lactate_diff + # I didn't include Lactate_min as low lactate levels are ideal/expected for a healthy person
                           MAP_max +
                           NISysABP_max +
                           NISysABP_min +
                           PaO2_diff +
                           PaO2_min,
                           data = icu_patients_df1, family = binomial)
summary(icu_glm_initial_2)
'
In this second model, the following are significant:
BUN_max                               0.0377269  0.0102347   3.686 0.000228 ***
SAPS1                                 0.1598306  0.0685511   2.332 0.019724 *  
Creatinine_max                       -0.2921214  0.1466432  -1.992 0.046365 *  
HCT_min                               0.1321537  0.0381673   3.462 0.000535 ***
MAP_diff                              0.0511316  0.0173223   2.952 0.003160 ** 
PaO2_max                             -0.0075585  0.0038625  -1.957 0.050362 .  
Urine_diff                           -0.0007600  0.0004575  -1.661 0.096640 .  
MAP_max                              -0.0440888  0.0162755  -2.709 0.006751 ** 
NISysABP_min                          0.0164209  0.0084823   1.936 0.052880 .  
PaO2_min                              0.0101487  0.0051102   1.986 0.047034 *  

AIC: 387.02 

Will look in other variables for PaO2, Urine, and blood pressure (systolic/diastolic/mean)

'


```



```{r}

# trial 3: trying other combinations

icu_glm_initial_3 <- glm(in_hospital_death ~ 
                           # patient descriptor variables
                           ICUType +
                           Age +
                           Gender +
                           Height +
                           BUN_max +
                           # clinical variables
                           SAPS1 + 
                           Creatinine_max +
                           DiasABP_diff + 
                           FiO2_max + 
                           GCS_min +
                           HCO3_diff +
                           HCT_min + 
                           HR_diff + 
                           K_diff + 
                           Lactate_max +
                           MAP_diff + 
                           Mg_diff + 
                           Na_diff +
                           NIDiasABP_diff + 
                           NIMAP_diff +
                           NISysABP_diff +
                           PaCO2_diff +
                           PaCO2_max + 
                           PaO2_max +
                           Platelets_min +
                           RespRate_diff + 
                           SaO2_diff + 
                           SysABP_diff +
                           Temp_diff +
                           TroponinI_max +
                           TroponinT_max +
                           Urine_diff +
                           WBC_diff +
                           Weight_max +
                           #EXPLORING OTHER VARIABLES
                           SOFA +
                           Lactate_diff + # I didn't include Lactate_min as low lactate levels are ideal/expected for a healthy person
                           MAP_max +
                           NISysABP_max +
                           NISysABP_min +
                           PaO2_diff +
                           PaO2_min +
                           #ADD ONS TO MODEL 3
                           PaO2_diff +
                           SysABP_min + #blood pressure variables
                           SysABP_max +
                           DiasABP_min +
                           DiasABP_max +
                           NIDiasABP_min +
                           NIDiasABP_max +
                           MAP_min +
                           Urine_min +
                           Urine_max,
                         data = icu_patients_df1, family = binomial)
summary(icu_glm_initial_3)

'
On third iteration, I can see some variables are strongly significant across trial models,
Though with varying degrees of significance, 
other variables of interest consistently come out in the models

BUN_max                               3.701e-02  1.142e-02   3.241 0.001192 ** 
SAPS1                                 1.270e-01  7.460e-02   1.702 0.088720 .  
Creatinine_max                       -5.115e-01  1.861e-01  -2.748 0.005988 ** 
HCT_min                               1.386e-01  4.042e-02   3.428 0.000608 ***
MAP_diff                              1.010e-01  3.232e-02   3.125 0.001777 ** 
PaO2_max                             -7.916e-03  4.186e-03  -1.891 0.058615 .  
Urine_diff                            1.511e-02  4.524e-03   3.340 0.000837 ***
MAP_max                              -9.012e-02  3.081e-02  -2.925 0.003447 ** 
NISysABP_min                          2.651e-02  1.287e-02   2.060 0.039400 *  
PaO2_min                              1.157e-02  5.345e-03   2.165 0.030403 *  
DiasABP_min                          -5.038e-02  3.004e-02  -1.677 0.093538 .  
MAP_min                               4.820e-02  2.712e-02   1.777 0.075488 .  
Urine_max                            -1.546e-02  4.410e-03  -3.505 0.000456 ***

AIC: 384.28

Patient descriptors didnt seem to have any significance so far, will cross out for next iteration.
If we group the clinical variables, we have markers for the following so far:
    blood pressure (MAP, SysABP)
    kidney function (BUN, urine, creatinine)
    blood O2 levels (PaO2)
    assessment tool/score (SAPS1)


Ive reviewed the data and would like to check a few more clinical variables which may be of interest/significance:
    glucose
    albumin
    cholesterol
    liver profile markers (AST, ALT, ALP)
    electrolytes (Na, Mg, K)


'

```

```{r}


# trial 4
icu_glm_initial_4 <- glm(in_hospital_death ~ 
                           # CLINICAL VARIABLES GROUPED FROM FIRST 3 TRIAL MODELS
                           # rapid assessment tool
                           SAPS1 +
                           # kidney markers
                           BUN_max +
                           Creatinine_max +
                           Urine_diff +
                           Urine_max +
                           # blood pressure variables
                           MAP_diff +
                           MAP_max +
                           MAP_min +
                           NISysABP_min +
                           DiasABP_min +
                           # blood levels
                           PaO2_max +
                           PaO2_min +
                           HCT_min +
                           # EXPLORING OTHER VARIABLES OF INTEREST
                           Glucose_diff +
                           Albumin_diff +
                           ALP_max +
                           ALT_max +
                           AST_max +
                           K_min +
                           Na_diff +
                           Mg_min +
                           Cholesterol_max,
                         data = icu_patients_df1, family = binomial)
summary(icu_glm_initial_4)


'
From trial model 4, none of the added variables of interest were significant. 

SAPS1            0.1302302  0.0268884   4.843 1.28e-06 ***
BUN_max          0.0275854  0.0061310   4.499 6.82e-06 ***
Creatinine_max  -0.3569116  0.1083592  -3.294 0.000988 ***
Urine_diff       0.0079677  0.0026051   3.059 0.002224 ** 
Urine_max       -0.0086481  0.0024959  -3.465 0.000530 ***
MAP_diff         0.0279541  0.0112775   2.479 0.013185 *  
MAP_max         -0.0251078  0.0106622  -2.355 0.018530 *  
MAP_min          0.0189800  0.0099203   1.913 0.055718 .  
DiasABP_min     -0.0250375  0.0125252  -1.999 0.045612 *  
PaO2_max        -0.0019022  0.0010563  -1.801 0.071739 .  
PaO2_min         0.0051129  0.0021546   2.373 0.017645 *  
HCT_min          0.0444069  0.0202447   2.194 0.028271 *  

AIC: 705.14

Ill try running these for trial model5
'
```



```{r}

# trial 5: preliminary model from the first 4 iterations
# variables trimeed down, chosen based on clinical background and statistical signifance
icu_glm_initial_5 <- glm(in_hospital_death ~ 
                           # CLINICAL VARIABLES GROUPED FROM FIRST 3 TRIAL MODELS
                           # rapid assessment tool
                           SAPS1 +
                           # kidney markers
                           BUN_max +
                           Creatinine_max +
                           Urine_diff +
                           Urine_max +
                           # blood pressure variables
                           MAP_diff +
                           MAP_max +
                           MAP_min +
                           NISysABP_min +
                           DiasABP_min +
                           # blood levels
                           PaO2_max +
                           PaO2_min +
                           HCT_min,
                         data = icu_patients_df1, family = binomial)
summary(icu_glm_initial_5)
```
Im not quite sure why, but the AIC has inflated to 700s in the latter models despite removing insignificant variables. 
Happy to start with these 13 variables for analysis.
I have grouped them (see code comments) to help with our decision making later in eliminating variables of the same category/context.


```{r}
# START OF GLM MODELLING
# "icu_glm_initial_5"
# this now contains the initial 13 variables for analysis and the starting point of modelling

```


```{r}
```


```{r}
```


```{r}
```
